---
title: "Lab 1"
author: "Taylor"
date: "1/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Objectives:

- basic review of data wrangling
- intro to spatial data in R with sf
- intro to shiny apps

1. Attach packages

```{r load_packages, include = FALSE}

library(tidyverse)
library(sf)
library(leaflet)
library(tmap)

```

2. Review: reading, exploring, wrangling and visualizing data

1. Read in the data
```{r data}

wine <- read_csv('wine_data.csv')

# Let's prioritize price and points to price ratio. And focus on Pinot Noir.

```

b. Wrangling
```{r wrangle}

wine_new <- wine %>% #command shift M
  select(country, province, winery, region_1, points, price) %>% 
    # select is for columns; filter is for rows
  rename(state = province) %>% 
  filter(state == 'California' | state == 'Oregon' | state == 'Washington') %>% 
    # single = sets equal; double == looks for a match
  mutate(ppd = points/price) %>% 
    # ppd = points per dollar
  arrange (-ppd) 
    # decreasing value

```

Let's say we want to summarize characteristics of wine based on states. We could create three separate data sets, but... we could also make R do it, which is way easier. SO:

c. Use group_by() and summarize() to find summary statistics by group.
```{r summary_stats}

wine_summary <- wine_new %>% 
  group_by(state) %>% 
    # this will not look ANY DIFFERENT -- the difference is just in R's brain! 
  summarize(
    mean_ppd = mean(ppd, na.rm = TRUE), #please remove any NA values
    max_ppd = max(ppd, na.rm = TRUE)
  )

```

MEAN: 
California: 2.428557
Oregon: 2.419863
Washington: 3.345340 <- WA is awesome! BUT 

MAX:
California: 12 <- max is in CA. Flipflop, baybee.
Oregon: 9
Washington: 7

Nice. BUT HOLD UP -- DON'T FORGET TO LOOK AT YOUR DATA!

d. A few GG Graphics.

First: histograms of price per dollar values for each state.
```{r histograms}

wine_hist <- ggplot(wine_new, aes(x = ppd)) +
  geom_histogram(aes(fill = state)) +
    # so that looks pretty good, but WA is too tiny, so let's split it up
  facet_wrap(~state, scales = 'free')
    # OK, so... splits it but still a bad scale for WA. Overide scales.

wine_hist
```

New graph: scatterplot graph of price v points, with color determined by state, adding a linear model trendline on top of it.

```{r scatterplot}

wine_points <- ggplot(wine_new, aes(x = price, y = points)) +
  geom_point(aes(color=state), alpha = 0.5) + 
    # POINTS are COLOR, BARS are FILL
  geom_smooth(color = 'black', method = 'lm', se = FALSE) +
    # Easy way to add linear model!!!! Look into this!
  scale_color_manual(values = c('royalblue','mediumorchid1','orange')) +
  theme_classic() + # look at different packages and themes
  facet_wrap(~state)

wine_points

```

2. California counties (spatial data) using SF

```{r sf_start}

ca_counties <- st_read(dsn = ".", layer = 'CA_Counties_TIGER2016')

# there are packages that pull data directly from census, etc.

# plot(ca_counties) 

## WOW COOL!

### figure out how to insert images in GitHub

# here's a simplified one --

ca_land <- ca_counties %>% 
  select(NAME, ALAND)

# NOW we are going to merge population income data.

ca_pop_inc <- read_csv('ca_pop_inc.csv') %>% 
  # OK so now we need to merge them, but counties are called NAME in ca_counties and COUNTY in 
  # ca_pop_inc. We gotta rename. REMEMBER it's NEW = OLD
  rename(NAME = COUNTY)

ca_df <- full_join(ca_land, ca_pop_inc) %>% 
  select(NAME, MedFamilyIncome) # but the geoms are sticky because SF is awesome.

```

d. Use geom_sf to make a map in ggplot

```{r mapping1}

ca_income <- ggplot(ca_df) +
  geom_sf(aes(fill = MedFamilyIncome)) +
  theme_classic()

ca_income
```

```{r leaflet}

# gotta change the data to match Leaflet

ca_df_transform <- st_transform(ca_df, crs = 4326)

leaflet(ca_df_transform) %>%
  addTiles() %>% 
  addPolygons()

```

